Analysis_v1.1_weather

### 2022-01-13 R/Analysis_v1.1_weather_2020.Rmd
Add 25ish site*year combination phenotypic variance-covariance matrices and visualization

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(zoo)
library(cowplot)
```

@@

@@

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(zoo)
library(lubridate)
library(cowplot)
```

@@

# ----
## load phe & wea data
```{r}
phe_ten_sites <- readRDS(here("data", "Phenotypes_2018_to_2021_10_sites.rds"))
wea_ten_sites <- readRDS(file = here("data", "Daily_weather_2018_to_2021_10_sites.rds"))

@@

# Phe correlation matrices

## load data
```{r}
gr_ten_sites <- readRDS(file = here("data", "Weather_related_phe_GR50_3yr.rds"))
fl_ten_sites <- readRDS(here("data", "Weather_related_phe_FL50_3yr.rds"))

### All site*years
Make phenotypic correlation matrices for greenup & flowering for these new years for all site*year combinations.

fl_wide <- fl_ten_sites %>%
  select(PLANT_ID:FL50, SUBPOP) %>%
  select(-GR50) %>%
  pivot_wider(names_from = c("manu_site", "Year"), values_from = "FL50")
gr_wide <- fl_ten_sites %>%
  select(PLANT_ID:GR50, SUBPOP) %>%
  pivot_wider(names_from = c("manu_site", "Year"), values_from = "GR50")

k=1
U_hyp_sitephe <- list()
library(rlist)
for(k in seq_along(subpop_v2)){

    cov_df <- fl_wide %>%
      ungroup() %>%
      filter(SUBPOP %in% subpop_v2[[k]])

    cor_phe <- cor(cov_df[,-(1:2)], use = "pairwise")
    # Set diagonal of this matrix = coefficient of variation within each garden.
    diag(cor_phe) <- cov_df %>%
      ungroup() %>%
      select(where(is.numeric)) %>%
      summarise(across(everything(), ~ sqrt(var(.x, na.rm = TRUE))/mean(.x, na.rm = TRUE))) %>%
      as_vector()

    U_hyp_sitephe <- list.prepend(U_hyp_sitephe, cor_phe)
    names(U_hyp_sitephe)[1] <- paste0("FL50", "_", names(subpop_v2)[k])
}

for(k in seq_along(subpop_v2)){

    cov_df <- gr_wide %>%
      ungroup() %>%
      filter(SUBPOP %in% subpop_v2[[k]])

    cor_phe <- cor(cov_df[,-(1:2)], use = "pairwise")
    # Set diagonal of this matrix = coefficient of variation within each garden.
    diag(cor_phe) <- cov_df %>%
      ungroup() %>%
      select(where(is.numeric)) %>%
      summarise(across(everything(), ~ sqrt(var(.x, na.rm = TRUE))/mean(.x, na.rm = TRUE))) %>%
      as_vector()

    U_hyp_sitephe <- list.prepend(U_hyp_sitephe, cor_phe)
    names(U_hyp_sitephe)[1] <- paste0("GR50", "_", names(subpop_v2)[k])
}

```{r}
plot_corr_fl_gulf <- U_hyp_sitephe$FL50_Gulf %>%
  as_tibble() %>%
  add_column(matrix_row = colnames(U_hyp_sitephe$FL50_Gulf), .before = TRUE) %>%
  pivot_longer(cols = -matrix_row, names_to = "matrix_col", values_to = "pvalue") %>%
  #filter(matrix_row != matrix_col) %>%
  arrange(pvalue) %>%
  ggplot(aes(x = matrix_col, y = matrix_row)) + 
  geom_tile(aes(fill = pvalue)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
      scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
                                      "white", "#27AD81FF", "#5DC863FF",
                                      "#FDE725FF"), limits = c(-1,1)) +
  labs(title = "Gulf Flowering P Corr All Site*Years", 
       subtitle = "Grey indicates NA values")

save_plot(plot = plot_corr_fl_gulf, filename = here("analysis", "hypothesis_matrices", "Weather_Uhyp_Corr_FL_Gulf.png"), base_height = 4.5, base_width = 5)

plot_corr_fl_midw <- U_hyp_sitephe$FL50_Midwest %>%
  as_tibble() %>%
  add_column(matrix_row = colnames(U_hyp_sitephe$FL50_Midwest), .before = TRUE) %>%
  pivot_longer(cols = -matrix_row, names_to = "matrix_col", values_to = "pvalue") %>%
  #filter(matrix_row != matrix_col) %>%
  arrange(pvalue) %>%
  ggplot(aes(x = matrix_col, y = matrix_row)) + 
  geom_tile(aes(fill = pvalue)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
      scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
                                      "white", "#27AD81FF", "#5DC863FF",
                                      "#FDE725FF"), limits = c(-1,1)) +
  labs(title = "Midwest Flowering P Corr All Site*Years", 
       subtitle = "Grey indicates NA values")

save_plot(plot = plot_corr_fl_midw, filename = here("analysis", "hypothesis_matrices", "Weather_Uhyp_Corr_FL_Midwest.png"), base_height = 4.5, base_width = 5)

plot_corr_fl_gam <- U_hyp_sitephe$FL50_Gulf_and_Midwest %>%
  as_tibble() %>%
  add_column(matrix_row = colnames(U_hyp_sitephe$FL50_Gulf_and_Midwest), .before = TRUE) %>%
  pivot_longer(cols = -matrix_row, names_to = "matrix_col", values_to = "pvalue") %>%
  #filter(matrix_row != matrix_col) %>%
  arrange(pvalue) %>%
  ggplot(aes(x = matrix_col, y = matrix_row)) + 
  geom_tile(aes(fill = pvalue)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
      scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
                                      "white", "#27AD81FF", "#5DC863FF",
                                      "#FDE725FF"), limits = c(-1,1)) +
  labs(title = "Gulf & Midwest Flowering P Corr All Site*Years", 
       subtitle = "Grey indicates NA values")

save_plot(plot = plot_corr_fl_gam, filename = here("analysis", "hypothesis_matrices", "Weather_Uhyp_Corr_FL_GAM.png"), base_height = 4.5, base_width = 5)

plot_corr_gr_gulf <- U_hyp_sitephe$GR50_Gulf %>%
  as_tibble() %>%
  add_column(matrix_row = colnames(U_hyp_sitephe$GR50_Gulf), .before = TRUE) %>%
  pivot_longer(cols = -matrix_row, names_to = "matrix_col", values_to = "pvalue") %>%
  #filter(matrix_row != matrix_col) %>%
  arrange(pvalue) %>%
  ggplot(aes(x = matrix_col, y = matrix_row)) + 
  geom_tile(aes(fill = pvalue)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
      scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
                                      "white", "#27AD81FF", "#5DC863FF",
                                      "#FDE725FF"), limits = c(-1,1)) +
  labs(title = "Gulf Green Up P Corr All Site*Years", 
       subtitle = "Grey indicates NA values")

save_plot(plot = plot_corr_gr_gulf, filename = here("analysis", "hypothesis_matrices", "Weather_Uhyp_Corr_GR_Gulf.png"), base_height = 4.5, base_width = 5)

plot_corr_gr_midw <- U_hyp_sitephe$GR50_Midwest %>%
  as_tibble() %>%
  add_column(matrix_row = colnames(U_hyp_sitephe$GR50_Midwest), .before = TRUE) %>%
  pivot_longer(cols = -matrix_row, names_to = "matrix_col", values_to = "pvalue") %>%
  #filter(matrix_row != matrix_col) %>%
  arrange(pvalue) %>%
  ggplot(aes(x = matrix_col, y = matrix_row)) + 
  geom_tile(aes(fill = pvalue)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
      scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
                                      "white", "#27AD81FF", "#5DC863FF",
                                      "#FDE725FF"), limits = c(-1,1)) +
  labs(title = "Midwest Green Up P Corr All Site*Years", 
       subtitle = "Grey indicates NA values")

save_plot(plot = plot_corr_gr_midw, filename = here("analysis", "hypothesis_matrices", "Weather_Uhyp_Corr_GR_Midwest.png"), base_height = 4.5, base_width = 5)

plot_corr_gr_gam <- U_hyp_sitephe$GR50_Gulf_and_Midwest %>%
  as_tibble() %>%
  add_column(matrix_row = colnames(U_hyp_sitephe$GR50_Gulf_and_Midwest), .before = TRUE) %>%
  pivot_longer(cols = -matrix_row, names_to = "matrix_col", values_to = "pvalue") %>%
  #filter(matrix_row != matrix_col) %>%
  arrange(pvalue) %>%
  ggplot(aes(x = matrix_col, y = matrix_row)) + 
  geom_tile(aes(fill = pvalue)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
      scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
                                      "white", "#27AD81FF", "#5DC863FF",
                                      "#FDE725FF"), limits = c(-1,1)) +
  labs(title = "Gulf & Midwest Green Up P Corr All Site*Years", 
       subtitle = "Grey indicates NA values")

save_plot(plot = plot_corr_gr_gam, filename = here("analysis", "hypothesis_matrices", "Weather_Uhyp_Corr_GR_GAM.png"), base_height = 4.5, base_width = 5)
```
